---
title: "Topic Model Workshop - Part 3 (STM)"
author: "Ryan Wesslen"
date: "April 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

## Load the data

Let's start by loading our packages and loading the dataset.

```{r}
library(quanteda); library(tidyverse); library(RColorBrewer); library(stm)

dataset <- read_csv("./articles-sample.csv")

#create corpus
myCorpus <- corpus(dataset$Abstract)
docvars(myCorpus, field = "Subject") <- ifelse(dataset$College=="Computing and Informatics","Computing","Social Science")
docvars(myCorpus, field = "Year") <- as.integer(dataset$Year)
```

## Create the dfm (pre-processing)

```{r}
stopWords <- c("can","use","uses","used","using","study","studies","first","second","third","also","across","results","result","resulted","may","however","one","two","three","four","five","among","well","within","many","related","i.e","e.g","find","finding","finds","found","increase","increases","increasing","increased","decreased","decrease","decreases","decreasing","propose","proposal","proposals","proposes","proposed","new","old","differ","differs","different","difference","differences","positive","negative","findings","reports","report","reported","state","states","article","articles","examines","examine","suggest","research","researches","researchers","need","needs","show","shows","association","associations","associated","discuss","discusses","discussed","will","likely","unlikely","paper","method","methods","methodology","compared","specifically","approach","impact","impacts","examine","examined","examines","includes","include","included","including","measure","measures","measured","analysis","analyze","analyses","complete","completes","completed","indicate","indicated","indicates","high","higher","low","lower","follow","follows","following","significant","significance","approach","approaches","approached","model","models","demonstrate","demonstrated","demonstrates","yet","best","worst","better","large","small","larger","smaller","several","few","much","less","given","via","long","short","often","years","along","whether","potential","significantly","influence","influenced","influences","develop","develops","developed","good","bad","based","p","group","groups","effect","affect","affects","effects","sample","samples","relationship","relationships","change","changes","m","k","conclusion","conclusions","present","presents")

dfm <- dfm(myCorpus,
           remove = c(stopwords("english"), stopWords),
           ngrams= 1L,
           stem = F,
           remove_numbers = TRUE, 
           remove_punct = TRUE,
           remove_symbols = TRUE)
```

## Structural Topic Model: Subject and Year

For this part, we're going to use the `stm` package to introduce two covariates into our model: field (Computing or Social Science) and year.

```{r}
library(stm)

# use quanteda converter to convert our Dfm
stmdfm <- convert(dfm, to = "stm", docvars = docvars(myCorpus))
```

Unlike the `topicmodels` packages, `stm` has built in features to help analysts reduce sparse terms (minDoc or minCount).

```{r fig.height=4}
plotRemoved(stmdfm$documents, lower.thresh = seq(1, 100, by = 10))

out <- prepDocuments(stmdfm$documents, stmdfm$vocab, stmdfm$meta, lower.thresh = 5)
```

Let's run a 40 topic model.

```{r}
k <- 40
load("./stmFit.RData")

#stmFit <- stm(out$documents, out$vocab, K = k, prevalence =~ s(Year) + Subject, 
#              max.em.its = 150, data = out$meta, init.type = "Spectral", seed = 300)

#save(stmFit, file = "./stmFit.RData")
```

Let's explore the topics.

```{r, fig.height=10, fig.width=8}
plot(stmFit, 
         type = "summary", 
         xlim = c(0,.16), 
         n = 5, 
         labeltype = "prob",
         main = "UNCC Research Topics", 
         text.cex = 0.8)
```

We can see almost identical topics -- this is a good sign. Our topics are "stable" across runs, even in this case after we added in two prevalent covariates (Year and Subject).

Let's save our topic information.

```{r}
topicNames <- labelTopics(stmFit, n = 5)
topic <- data.frame(
  TopicNumber = 1:k,
  TopicProportions = colMeans(stmFit$theta))
```

## Exploring the effects of the covariates: Subject

Next, we want to explore the effect of the covariates on the topic proportions.

First, we'll need to use the `estimateEffect` function to estimate this effect.

```{r}
prep <- estimateEffect(1:k ~ Subject + s(Year), stmFit, meta = out$meta, uncertainty = "Global")
```

We can then use the `plot.estimateEffect` function to compare the effect of the "Subject" field (Computing or Social Science binary flag) on topic proportions (likelihood of the topic).

```{r fig.height = 10, fig.width = 8}
Result <- plot(prep, "Subject", method = "difference", 
                              cov.value1 = "Social Science", cov.value2 = "Computing", 
                              verbose.labels = F, 
                              ylab = "Expected Difference in Topic Probability by Subject (with 95% CI)", 
                              xlab = "More Likely Computing                           Not Significant                       More Likely Social Science",
                              main = "Effect of Subject on Topic Prevelance for UNCC Research",
                              xlim = c(-0.1,0.1))
```

Let's redo this plot but rank the topics.

```{r fig.height = 10, fig.width = 8}
# order based on Expected Topic Proportion
rank = order(unlist(Result$means))
topicRnk <- topic[rank,]

plot(prep, "Subject", method = "difference", 
                              cov.value1 = "Social Science", cov.value2 = "Computing", 
                              verbose.labels = F, 
                              topics = topicRnk$TopicNumber,
                              #labeltype = "custom",
                              #custom.labels  = apply(topicNames$prob, 1, function(x) paste0(x, collapse = " + ")),
                              ylab = "Expected Difference in Topic Probability by Subject (with 95% CI)", 
                              xlab = "More Likely Computing                           Not Significant                       More Likely Social Science",
                              main = "Effect of Subject on Topic Prevelance for UNCC Research",
                              xlim = c(-0.1,0.1))
```

## Effect of Time

```{r}
# time
par(mfrow = c(1,1),mar = c(4,4,2,2))
i <- c(9,18)
  plot(prep, "Year", method = "continuous", topics = i, 
                      main = "Topics 9 and 18 by Year",
                      printlegend = T, ylab = "Exp. Topic Prob", 
                      xlab = "Year", ylim = c(-0.01,0.16)
                      )

```

